### 2013-03-28 doc/sosp13/evaluation.tex

\subsection{Microbenchmark Results}
\subsubsection{Random Access}
\TODO{Random access feed forward results on Grappa.  optional: Results we measured for XMT.  Cite MPI results}
\subsubsection{Latency Tolerance}
\TODO{Simple ping test results -- eg, MPI ping, not the full blown aggregation ping test.  Random access blocking results on Grappa.}
\subsubsection{Scheduling and Robustness}
\TODO{Yield test results:  latency \& bandwidth.  Discussion of implications of zillions of contexts on robustness of latency tolerance preshadowing ~\ref{sec:scaling}}

>>>>

\subsection{Microbenchmark Results}

\subsubsection{Random Access}
\TODO{Random access feed forward results on Grappa.  optional: Results we measured for XMT.  Cite MPI results}
\subsubsection{Latency Tolerance}
\TODO{Simple ping test results -- eg, MPI ping, not the full blown aggregation ping test.  Random access blocking results on Grappa.}
\subsubsection{Scheduling and Robustness}

\TODO{Summary of what to write please expand: Reference above plots. Bandwidth of single socket is
    270Mcacheline/s. Each context is 4 cachelines: 1 for worker struct
    and 3 stack cachelines. We must read and write every context, so 8
    cacheline transfers per context switch. This asymptotically
    approaches 33.75Mcontexts/s as you increase the number of contexts
    (fewer and fewer in L3). Note that only 4+ cores reaches full rate
    because these westmere chips are balanced to not achieve full off-chip bandwidth until 4
    cores--bdmyers}

\TODO{Yield test results:  latency \& bandwidth.  Discussion of implications of zillions of contexts on robustness of latency tolerance preshadowing ~\ref{sec:scaling}}

### 2013-03-28 doc/sosp13/method.tex

To study the Grappa system we implemented it in C++ for the Linux operating system.  We ported a number of benchmarks to Grappa as well as collected and optimized a set of comparison benchmarks for alternative approaches.  We run Grappa on a cluster of AMD Interlagos processors. Nodes have 32-core 2.1-GHz processors, 64GB of memory, and 40Gb Mellanox ConnectX-2 infiniband network cards.  Nodes are connected via a QLogic infiniband switch.  We use this cluster to evaluate Grappa and native MPI-only performance.  In addition to a mass-market cluster of machines we compare to the MTA using a 128-node Cray XMT (3rd generation MTA). Each node consists of a 500-MHz MTA Threadstorm multithreaded processor that supports 128~streams. The machine uses Cray's proprietary SeaStar2 interconnection network.  To evaluate performance we use a variety of benchmarks:

\paragraph{Context-switch:} This simple microbenchmark is used to explore context switch overhead.  A configurable number of workers are used to increment the values in a large array.  This benchmark stresses the scheduler inside of Grappa.

\paragraph{GUPS:} This is a faithful implementation of the giga-updates-per-second benchmark.  Read-modify-write updates are dispatched at random to a large array.  This benchmark stresses the networking layer of Grappa separately from the scheduler, because only a single worker is used per system node.

\paragraph{Unbalanced tree search in-memory (UTS-Mem):} Unbalanced Tree Search (UTS) is a benchmark for evaluating the programmability and performance of systems for parallel applications that require dynamic load balancing~\cite{UTS}. It involves traversing an unbalanced implicit tree: at each vertex, its number of children is sampled from some probability distribution, and this number of new nodes are added to a work queue to be visited. While this benchmark captures irregular, dynamic \emph{computation,} we actually want to evaluate performance of algorithms with irregular \emph{memory\/} access patterns. Thus we augment UTS by using the existing traversal code to create a large tree in memory, and then we traverse the in-memory tree. We call this benchmark UTS-Mem, and the timed portion is this traversal of the in-memory tree. This in-memory traversal has no knowledge of the tree structure beforehand.

\paragraph{Breadth-first-search (BFS):} This is the primary kernel for the Graph500 benchmark and is what currently determines the ranking of machines on the Graph500 list~\cite{graph500list}. As a whole, the Graph500 benchmark suite is designed to bring the focus of system design on data-intensive workloads, particularly large-scale graph analysis problems, that are important among cybersecurity, informatics, and network understanding workloads. The BFS benchmark builds a search tree containing parent nodes for each traversed vertex during the search.  While this is a relatively simple problem to solve, it exercises the random-access and fine-grained synchronization capabilities of a system as well as being a primitive in many other graph algorithms. Performance is measured in \emph{traversed edges per second\/} (TEPS), where the number of edges is the edges making up the generated BFS tree. One of the reference implementations of Graph500 BFS is for the XMT; this code fails to scale past 16 XMT processors because it does not expose enough parallelism, so we modified the code to use a recursive loop decomposition similar to Grappa's. We compare this modified version against a straightforward Grappa implementation. We do not employ algorithmic improvements, though there are many~\cite{Beamer:Graph500,Yoo:FixedPointGraph500}.

\paragraph{IntSort:}  Write me Brandon Holt!

The metric we use is algorithmic time, which means startup and loading of the data structure (from disk) is not included in the measurement.  Grappa collects statistics about application behavior (packets sent, context switches, etc) and these are discussed where appropriate.  In addition we use a profiled version of Grappa to report component runtimes, but this profiled version is only used to collect this data and a fully optimized version is used for the bulk of the presented results.

>>>>

We implemented the \Grappa in C++ for the Linux operating system. The core
runtime system system is about 15K lines of code. We ported a number of
benchmarks to \Grappa as well as collected and optimized a set of comparison
benchmarks for XMT, MPI and UPC~\cite{UPC}. We run the \Grappa, MPI and UPC
experiments on a cluster of AMD Interlagos processors. Nodes have 32 2.1-GHz
cores in two sockets, 64GB of memory, and 40Gb Mellanox ConnectX-2 InfiniBand
network cards. Nodes are connected via a QLogic InfiniBand switch. We also
compare \Grappa to a 128-node Cray XMT (3rd generation MTA). Each node
consists of a 500-MHz MTA Threadstorm multithreaded processor that supports
128~streams. The machine uses Cray's proprietary SeaStar2 interconnection
network.

We use a variety of benchmarks:

\noindent{\bf Context-switch:} This simple microbenchmark is used to explore context switch overhead.  A configurable number of workers are used to increment the values in a large array.  This benchmark stresses the scheduler inside of Grappa.

\nonindent{\bf GUPS:} This is a faithful implementation of the
giga-updates-per-second benchmark. Read-modify-write updates are dispatched at
random to a large array. This benchmark stresses the networking layer of
Grappa separately from the scheduler, because only a single worker is used per
system node.

\nonindent{\bf Unbalanced tree search in-memory (UTS-Mem):} Unbalanced Tree Search
(UTS) is a benchmark for evaluating the programmability and performance of
systems for parallel applications that require dynamic load
balancing~\cite{UTS}. It involves traversing an unbalanced implicit tree: at
each vertex, its number of children is sampled from some probability
distribution, and this number of new nodes are added to a work queue to be
visited. While this benchmark captures irregular, dynamic \emph{computation,}
we actually want to evaluate performance of algorithms with irregular
\emph{memory\/} access patterns. Thus we augment UTS by using the existing
traversal code to create a large tree in memory, and then we traverse the
in-memory tree. We call this benchmark UTS-Mem, and the timed portion is this
traversal of the in-memory tree. This in-memory traversal has no knowledge of
the tree structure beforehand.

\nonindent{\bf Breadth-first-search (BFS):} This is the primary kernel for the
Graph500 benchmark and is what currently determines the ranking of machines on
the Graph500 list~\cite{graph500list}. As a whole, the Graph500 benchmark
suite is designed to bring the focus of system design on data-intensive
workloads, particularly large-scale graph analysis problems, that are
important among cybersecurity, informatics, and network understanding
workloads. The BFS benchmark builds a search tree containing parent nodes for
each traversed vertex during the search. While this is a relatively simple
problem to solve, it exercises the random-access and fine-grained
synchronization capabilities of a system as well as being a primitive in many
other graph algorithms. Performance is measured in \emph{traversed edges per
second\/} (TEPS), where the number of edges is the edges making up the
generated BFS tree. One of the reference implementations of Graph500 BFS is
for the XMT; this code fails to scale past 16 XMT processors because it does
not expose enough parallelism, so we modified the code to use a recursive loop
decomposition similar to Grappa's. We compare this modified version against a
straightforward Grappa implementation. We do not employ algorithmic
improvements, though there are
many~\cite{Beamer:Graph500,Yoo:FixedPointGraph500}.

\nonindent{\bf IntSort:} \TODO{Write me Brandon Holt!}

\nonindent{\bf PageRank:} \TODO{Write me Brandon Myers!}

The metric we use is algorithmic time, which means startup and loading of the
data structure (from disk) is not included in the measurement. \Grappa
collects statistics about application behavior (packets sent, context
switches, etc) and these are discussed where appropriate. In addition we use a
profiled version of \Grappa to report component runtimes, but this profiled
version is only used to collect this data and a fully optimized version is
used for the bulk of the presented results.
